\bf A simple algorithm for computing approximate Nash equilibria in two-person zero-sum games with imcomplete information

### 2015-06-07 nips2015.tex
getting down to business (WINE conf is coming)

\documentclass[envcountsame]{llncs} % For LaTeX2e
%\usepackage{mathpazo}
%\usepackage{nips14submit_e,times}
%\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath,amsfonts}%,amsthm}
\usepackage{bbm}
\usepackage{algorithm,algorithmicx}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}

\usepackage{algpseudocode}
\usepackage{pifont}

%\usepackage[section]{placeins}
\usepackage[FIGTOPCAP]{subfigure}

\def\B#1{\bm{#1}}
%\def\B#1{\mathbf{#1}}
\def\trans{\mathsf{T}}

%\renewcommand{\labelitemi}{--}

%\newtheorem{theorem}{Theorem} \newtheorem{lemma}[theorem]{Lemma}

%\newtheorem{proposition}[theorem]{Proposition}
%% \newtheorem{corollary}[theorem]{Corollary}
%% \newtheorem{definition}[theorem]{Definition}
%\newtheorem{remark}{Remark}
\newtheorem{fact}{Fact}

\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\proj}{proj}
\DeclareMathOperator{\prox}{prox}

@@

\begin{figure}[!htpb]
  \subfigure[Matrix game on simplices]{
    \includegraphics[width=.33\linewidth]{simplex_dgap.pdf}
  }
  \hspace{-2em}
  \subfigure[Simplified Poker]{
    \includegraphics[width=.33\linewidth]{SimplifiedPoker_dgap.pdf}
  }
  \hspace{-2em}
  \subfigure[Kuhn 3-card Poker]{
    \includegraphics[width=.33\linewidth]{Kuhn3112_dgap.pdf}
  }
  \vspace{-1em}
  \hspace{-1em}
  \includegraphics[width=.34\linewidth]{simplex_NE.pdf}
  \hspace{-1em}
  \includegraphics[width=.33\linewidth]{SimplifiedPoker_NE.pdf}
  \hspace{-1em}
  \includegraphics[width=.33\linewidth]{Kuhn3112_NE.pdf}
  \caption{Convergence curves of Algorithm
    \ref{Tab:algo}. \textbf{Top row}: convergence in terms of ergodic
    primal-dual gap. \textbf{Bottom row}: Convergence in terms of
    value of game.}
  \label{Tab:dgap_curve}
\end{figure}

\paragraph{\textbf{Simulated game: matrix game on simplexes.}}
Any matrix $A \in \mathbb{R}^{n_1 \times n_2}$ specifies a
\textit{matrix game}. The strategy profile of player $k$ is a simplex;
this simplex can be written in the form \eqref{eq:polyhedron} by
taking $E_k := [1, 1, ..., 1] \in \mathbb{R}^{1 \times n_k}$ and $e_k
= 1$. Thus every matrix game on simplexes can be seen as a sequential
game. Now, as in \textbf{nesterov, c-p, etc.}, we generate a $1000
\times 1000$ random matrix whose entries are uniformly identically
distributed in the closed interval $[-1, 1]$. See Figure \ref{Tab:sim_dgap_curve}...

\paragraph{\textbf{Real-world game: Kuhn 3-card Poker.}}
The sequence-form representation of this simplified Poker variant
\cite{kuhn} is given by (not showing zero entries)\\

$E_1 = \left[\begin{array}{ccccccccccccc}
1 &   &   &   &   &   &   &   &   &   &   &   &  \\
-1 &   &   &   &   &   &   &   &   & 1 &   &   & 1\\
-1 & 1 &   &   & 1 &   &   &   &   &   &   &   &  \\
-1 &   &   &   &   & 1 &   &   & 1 &   &   &   &  \\
  & -1 & 1 & 1 &   &   &   &   &   &   &   &   &  \\
  &   &   &   &   & -1 & 1 & 1 &   &   &   &   &  \\
  &   &   &   &   &   &   &   &   & -1 & 1 & 1 &  
\end{array}\right] \in \mathbb{R}^{7 \times 13}$,
$E_2 = \left[\begin{array}{ccccccccccccc}
1 &   &   &   &   &   &   &   &   &   &   &   &  \\
-1 &   &   &   &   &   &   & 1 & 1 &   &   &   &  \\
-1 &   &   &   &   &   &   &   &   & 1 & 1 &   &  \\
-1 &   &   &   &   & 1 & 1 &   &   &   &   &   &  \\
-1 &   &   &   &   &   &   &   &   &   &   & 1 & 1\\
-1 & 1 & 1 &   &   &   &   &   &   &   &   &   &  \\
-1 &   &   & 1 & 1 &   &   &   &   &   &   &   &  
\end{array}\right] \in \mathbb{R}^{7 \times 13}$, $e_1 = e_2 = [1, 0,
  0, 0, 0, 0, 0]^T \in \mathbb{R}^7$, and $A =
\\\left[\begin{array}{ccccccccccccc}

  &   &   &   &   &   &   &   &   &   &   &   &  \\
  &   &   &   &   &   &   & -1 / 6 &   &   &   & -1 / 6 &  \\
  &   &   &   &   &   &   &   & -1 / 6 &   &   &   & -1 / 6\\
  &   &   &   &   &   &   &   & -1 / 3 &   &   &   & -1 / 3\\
  &   &   &   &   & 1 / 6 & -1 / 3 &   &   & 1 / 6 & -1 / 3 &   &  \\
  &   &   & 1 / 6 &   &   &   &   &   &   &   & -1 / 6 &  \\
  &   &   &   & -1 / 6 &   &   &   &   &   &   &   & -1 / 6\\
  &   &   &   & 1 / 3 &   &   &   &   &   &   &   & -1 / 3\\
  & 1 / 6 & 1 / 3 &   &   &   &   &   &   & 1 / 6 & -1 / 3 &   &  \\
  &   &   & 1 / 6 &   &   &   & 1 / 6 &   &   &   &   &  \\
  &   &   &   & -1 / 6 &   &   &   & -1 / 6 &   &   &   &  \\
  &   &   &   & 1 / 3 &   &   &   & 1 / 3 &   &   &   &  \\
  & 1 / 6 & 1 / 3 &   &   & 1 / 6 & 1 / 3 &   &   &   &   &   &  
\end{array}\right] \in \mathbb{R}^{13 \times 13}$.\\
The pair $(x^*, y^*) \in \mathbb{R}^{13} \times \mathbb{R}^{13}$ of
realization plans given by

@@

\begin{eqnarray*}
  \begin{split}
    E_2y^* - e_2 = [&-7.04 \times 10^{-7}, 2.27 \times 10^{-6}, -3.29
      \times 10^{-6}, -1.50 \times 10^{-6},\\
      &2.92 \times 10^{-6}, -4.97 \times 10^{-7}, -5.85 \times
      10^{-7}]^T
    \end{split}
\end{eqnarray*}
Finally, one checks that
\begin{eqnarray*}
  {x^*}^TAy^* = -0.055593685705289997,
\end{eqnarray*}
 which agrees to 4 d.p with the value of $-1 / 18$ computed
 analytically by H. W. Kuhn in his 1950 paper \cite{kuhn}. The
 evolution of the dual gap and the expected value of the game across
 iterations is shown in Figure \ref{Tab:dgap_curve}. These figures
 validate the theoritically established $\mathcal{O}(1/\epsilon)$
 convergence rate of the Algorithm.

\section{Conclusion}
Making use of the sequence-form representation
\cite{koller1992complexity,von1996efficient,vonequilibrium}, we have
deviced a primal-dual algorithm for computing Nash equilibria in
two-person zero-sum sequential games with imcomplete information (like
Texas Hold'em, etc.). Our algorithm is simple to implement, with a
very low constant cost per iteration, and enjoys a rigorous
convergence theory with a proven $\mathcal{O}(1/\epsilon)$ convergence
in terms of basic operations (matvec products, clipping, etc.), to a
Nash $\epsilon$-equilibrium of the game. The core of our contribution
(summarized in Theorems \ref{thm:pd} and \ref{thm:conv}) was to show
that a reformulation of the Nash equilibrium problem for sequential
games with imcomplete information lends the problem directly
accessible to primal-dual framework
\cite{chambolle2010,chambolle2014ergodic}.

@@

\paragraph{Acknowledgments:} ...
\small
\bibliographystyle{splncs03}
\bibliography{bib}

>>>>

\documentclass[envcountsame]{llncs}
\usepackage{url}
\usepackage{amsmath,amsfonts}
\usepackage{bbm}
\usepackage{algorithm,algorithmicx}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}
\usepackage{algpseudocode}
\usepackage{pifont}
\usepackage[FIGTOPCAP]{subfigure}

\newtheorem{fact}{Fact}
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\proj}{proj}
\DeclareMathOperator{\prox}{prox}

@@

\begin{figure}[!htpb]
  \subfigure[Matrix game on simplices]{
    \includegraphics[width=.33\linewidth]{simplex_dgap.pdf}
  }
  \hspace{-2em}
  \subfigure[Toy Poker]{
    \includegraphics[width=.33\linewidth]{SimplifiedPoker_dgap.pdf}
  }
  \hspace{-2em}
  \subfigure[Kuhn 3-card Poker]{
    \includegraphics[width=.33\linewidth]{Kuhn3112_dgap.pdf}
  }
  \vspace{-1em}
  \hspace{-1em}
  \includegraphics[width=.34\linewidth]{simplex_NE.pdf}
  \hspace{-1em}
  \includegraphics[width=.33\linewidth]{SimplifiedPoker_NE.pdf}
  \hspace{-1em}
  \includegraphics[width=.33\linewidth]{Kuhn3112_NE.pdf}
  \caption{Convergence curves of Algorithm
    \ref{Tab:algo}. \textbf{Top row}: Evolution of ergodic
    primal-dual gap. \textbf{Bottom row}: Evolution of
    value of game.}
  \label{Tab:dgap_curve}
\end{figure}

\paragraph{\textbf{Matrix game on simplexes.}}
Any matrix $A \in \mathbb{R}^{n_1 \times n_2}$ specifies a
\textit{matrix game}. The strategy profile of player $k$ is a simplex;
this simplex can be written in the form \eqref{eq:polyhedron} by
taking $E_k := [1, 1, ..., 1] \in \mathbb{R}^{1 \times n_k}$ and $e_k
= 1$. Thus every matrix game on simplexes can be seen as a sequential
game. Now, as in \textbf{nesterov, c-p, etc.}, we generate a $1000
\times 1000$ random matrix whose entries are uniformly identically
distributed in the closed interval $[-1, 1]$.

\paragraph{\textbf{Toy Poker game.}}
The has $n_1 = n_2 = 5$ and $l_1 = l_2 = 3$. Its sequence-form
representation is given by (showing only nonzero entries):\\

 $E_1(0,0) = E_1(1,1) =
E_1(1,2) = E_1(2,3) = E_1(2,4) = \textbf{1}$,
$E_1(1,0) = E_1(2,0) = \textbf{-1}$;\\

$E_2(0,0) = E_2(1,1) = E_2(1,2) = E_2(2,3) = E_2(2,4) = \textbf{1}$,
$E_2(1,0) = E_2(2,0) = \textbf{-1}$; and\\

$A(2,0) = A(4,0) = \textbf{-0.5}$, $A(1,3) = \textbf{1}$, $A(3,1) =
\textbf{-1}$, $A(1,2) = A(1,4) = A(3,2) = A(3,4) = \textbf{0.25}$.

\paragraph{\textbf{Kuhn 3-card Poker.}}
The game has $n_1 = n_2 = 13$ and $l_1 = l_2 = 7$, and its
sequence-form is given by:\\

$E_1(0,0) = E_1(1,9) = E_1(1,12) = E_1(2,1) = E_1(2,4) = E_1(3,5) =
E_1(3,8) = E_1(4,2) = E_1(4,3) = E_1(5,6) = E_1(5,7) = E_1(6,10) =
E_1(6,11) = \textbf{1}$, $E_1(1,0) = E_1(2,0) = E_1(3,0) = E_1(4,1) =
E_1(5,5) = E_1(6,9) = \textbf{-1}$;\\

$E_2(0,0) = E_2(1,7) = E_2(1,8) = E_2(2,9) = E_2(2,10) = E_2(3,5) =
E_2(3,6) = E_2(4,11) = E_2(4,12) = E_2(5,1) = E_2(5,2) = E_2(6,3) =
E_2(6,4) = \textbf{1}$, $E_2(1,0) = E_2(2,0) = E_2(3,0) = E_2(4,0) =
E_2(5,0) = E_2(6,0) = \textbf{-1}$; and\\

$A(3,8) = A(3,12) = A(4,6) = A(4,10) = A(7,12) = A(8,10) =
\textbf{-0.333333}$, $A(1,7) = A(1,11) = A(2,8) = A(2,12) = A(5,11) =
A(6,4) = A(6,12) = A(10,4) = A(10,8) = \textbf{-0.166667}$, $A(7,4) =
A(8,2) = A(11,4) = A(11,8) = A(12,2) = A(12,6) = \textbf{0.333333}$,
$A(4,5) = A(4,9) = A(5,3) = A(8,1) = A(8,9) = A(9,3) = A(9,7) =
A(12,1) = A(12,5) = \textbf{0.166667}$.\\

The pair $(x^*, y^*) \in \mathbb{R}^{13} \times \mathbb{R}^{13}$ of
realization plans given by

@@

\begin{eqnarray*}
  \begin{split}
    E_2y^* - e_2 = [&-7.04 \times 10^{-7}, 2.27 \times 10^{-6}, -3.29
      \times 10^{-6}, -1.50 \times 10^{-6},\\
      &2.92 \times 10^{-6}, -4.97 \times 10^{-7}, -5.85 \times
      10^{-7}]^T
    \end{split}
\end{eqnarray*}
Finally, one checks that
\begin{eqnarray*}
  {x^*}^TAy^* = \textbf{-0.05555},
\end{eqnarray*}
 which agrees to 5 d.p with the value of $-1 / 18$ computed
 analytically by H. W. Kuhn in his 1950 paper \cite{kuhn}. The
 evolution of the dual gap and the expected value of the game across
 iterations is shown in Figure \ref{Tab:dgap_curve}.

\section{Conclusion}
Making use of the sequence-form representation
\cite{koller1992complexity,von1996efficient,vonequilibrium}, we have
deviced a primal-dual algorithm for computing Nash equilibria in
two-person zero-sum sequential games with imcomplete information (like
Texas Hold'em, etc.). Our algorithm is simple to implement, with a
very low constant cost per iteration, and enjoys a rigorous
convergence theory with a proven $\mathcal{O}(1/min(\rho,\epsilon))$ convergence
in terms of basic operations (matvec products, clipping, etc.), to a
Nash $(\rho,\epsilon$-equilibrium of the game.

@@

\small
\bibliographystyle{splncs03}
\bibliography{bib}
